---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# jtracer

<!-- badges: start -->
<!-- badges: end -->

jtracer provides an R interface to jTRACE, a re-implementation of the **TRACE** model of spoken word recognition (McClelland & Elman, [1986](https://www.sciencedirect.com/science/article/pii/0010028586900150)) created by Strauss, Harris & Magnusson ([2007](https://magnuson.psy.uconn.edu/jtrace/)). Using the functions in this package you can generate the files you need to perform simulations in jTRACE.

## Installation

You can install it from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("bilingual-project/jtracer")
```

## Installing and launching jTRACE

First, you need to download jTRACE. You do it from R with:

```r
jtracer::jtrace_install()
```
This will download the jTRACE folder from [https://magnuson.psy.uconn.edu/jtrace/](https://magnuson.psy.uconn.edu/jtrace/) into your home directory (normally Documents) as `.jtrace`. You won't have to open this folder to use jtracer. This function will also check if a sufficiently recent version of Java is up and running (>1.4). If not, it will prompt you to do it. 

Once an appropriate version of Java is up and running, and jTRACE has been installed, you will be able to launch jTRACE with:

```r
jtracer::jtrace_launch()
```

## Extracting a lexicon from jTRACE

jTRACE has multiple lexica available to perform simulations. By default, it offers:

* biglex_lr
* biglex_lr_^a
* biglex901
* context_lex
* empty
* initial_lexicon
* sevenlex
* slex
* slex_pairs

You can import a lexicon from jTRACE as a data frame with `jtrace_get_lexicon`. For example:

```{r lexicon, echo=TRUE}
lex <- jtracer::jtrace_get_lexicon(lexicon = "slex")
head(lex)
```




