---
title: "Creating a language (phoneme inventory)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating a language (phoneme inventory)}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r prepare}
library(jtracer)
jtrace_install()
```

We can create new languages by creating a set of phonemes and scoring them across the seven feature dimensions the TRACE model considers. The `jtrace_create_language` takes care of generating a file jTRACE can read and run simulations with.

The `jtrace_create_language` takes, at least, a character value indicating the name you want to give your new language, a vector or phonemes (in jTRACE notation) of length M, and a M x 7 matrix or data frame with the scores (0-9) of each phoneme (rows) in each of the feature dimensions (columns). If the feature data frame includes row names with valid phoneme symbols, you can omit the `phonemes` argument. Optionally, you can also provide duration values (0-1) for each phoneme in each feature as a M x 7 matrix with same dimensions as the one provided in the `features` argument, and/or a M x M matrix stating allophonic relations between phonemes (with logical values, TRUE/FALSE). See the documentation for more details (`?jtrace_create_language`).

# Generate a phoneme inventory

First, we create a character vector with the phoneme symbols:

```{r phonemes}
p <- c("-", "a", "s", "d", "f", "g", "c") 
```

Then, we create a the features matrix (please note that scores are assigned randomly, and this example is not supposed to reflect a real-life case):

```{r features}
f <- data.frame(
  bur = c(9, 6, 4, 3, 1, 1, 2),
  voi = c(7, 4, 3, 3, 3, 3, 4),
  con = c(8, 2, 4, 2, 5, 5, 6),
  grd = c(4, 6, 1, 4, 6, 8, 6),
  dif = c(6, 3, 2, 6, 6, 6, 7),
  voc = c(3, 8, 1, 6, 6, 7, 4),
  pow = c(6, 4, 1, 6, 1, 1, 5)
)
```

Of course, it can be cumbersome to create a data frame like this. An alternative would be to create a table in an Excel spreadsheet and importing it to your R session or copy-pasting it into your console using the "Paste as tribble" adding from the [datapasta](https://github.com/MilesMcBain/datapasta) R package. 

We are all set up now!

```{r language_create}
jtrace_create_language(
  language_name = "my_language",
  phonemes = p,
  features = f
)
```

