---
title: "Extracting a language"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting a language}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Extracting a language

jTRACE includes a default language, which consists on a matrix with the scores of a series of phonemes across the seven feature dimensions considered by the TRACE model, a matrix with the duration score of each phoneme in each feature, and a matrix of logical values indicating the allophonic relations between phonemes. You can extract this language by running:

```{r language_get}
library(jtracer)
default_language <- jtrace_get_language(language_name = "default")
```

This function will return a named list, containing the feature scores, duration scalars, and allophonic relations between phonemes:

```{r language_features}
default_language$features
```

```{r language_duration}
default_language$duration_scalar
```


```{r language_allophones}
default_language$allophonic_relations
```