---
title: "Phoneme inventory"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Phoneme inventory}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(jtracer)
jtrace_install()
```

At the lowest level of the TRACE model's architecture lie the acoustic features associated with the phonemes that compose words in the lexicon. jTRACE considers seven feature dimensions (see `?phonemes`):  power (`pow`), vocalic (`vow`), diffusiveness (`dif`), acuteness (`acu`), consonantal (`con`), voicing (`voi`), and burst (`bur`). These are the original dimensions used by McClelland and Elman (1986) in their original implementation of TRACE. Prior to the simulation, the model is introduced to a set of familiar phonemes, mimicking its familiarity with a language. Each of these phonemes are coded by assigning them a 1-9 score in each of the feature dimensions. McClelland and Elman's original phoneme set was later extended by Mayor and Plunkett (2014).


```{r matrix}
library(jtracer)
data("phonemes")
Encoding(phonemes[["ipa"]]) <- "Latin-1"
Encoding(phonemes[["trace"]]) <- "Latin-1"

head(phonemes)

```

## Non-English phonemes

Both versions were explicitly designed to fit the English phonology. We have coded new phonemes, not restricted to the English phonology, in the `phonemes` data frame included in this package. This way, it is possible to run simlation on a model that is familiar to phonologies different than the English one, such as Spanish and Catalan. The dataframe includes three logical variables indicating whether a given phoneme is present in English (`is_english`), Spanish (`is_spanish`), or Catalan (`is_catalan`), respectively.

### English phoneme set

```{r list_english, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(gt)
library(dplyr)

phonemes %>%
  filter(is_english) %>% 
  select(-starts_with("is_")) %>% 
  rename_at(vars(pow:bur), toupper) %>% 
  gt(rowname_col = "id", groupname_col = "type") %>% 
  tab_spanner(md("**Acoustic features**"), columns = 6:12) %>% 
  cols_label(
    ipa = md("**IPA symbol**"),
    trace = md("**jTRACE symbol**"),
    description = ""
  ) %>% 
  data_color(
    columns = 6:12,
    colors = scales::col_numeric(
      palette = c(
        "white", "orange"), 
      domain = c(0, 9)
    )
  )
```

### Spanish phoneme set

```{r list_spanish, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
phonemes %>%
  filter(is_spanish) %>% 
  select(-starts_with("is_")) %>% 
  rename_at(vars(pow:bur), toupper) %>% 
  gt(rowname_col = "id", groupname_col = "type") %>% 
  tab_spanner(md("**Acoustic features**"), columns = 6:12) %>% 
  cols_label(
    ipa = md("**IPA symbol**"),
    trace = md("**jTRACE symbol**"),
    description = ""
  ) %>% 
  data_color(
    columns = 6:12,
    colors = scales::col_numeric(
      palette = c(
        "white", "orange"), 
      domain = c(0, 9)
    )
  )
```

### Catalan phoneme set

```{r list_catalan, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
phonemes %>%
  filter(is_catalan) %>% 
  select(-starts_with("is_")) %>% 
  rename_at(vars(pow:bur), toupper) %>% 
  gt(rowname_col = "id", groupname_col = "type") %>% 
  tab_spanner(md("**Acoustic features**"), columns = 6:12) %>% 
  cols_label(
    ipa = md("**IPA symbol**"),
    trace = md("**jTRACE symbol**"),
    description = ""
  ) %>% 
  data_color(
    columns = 6:12,
    colors = scales::col_numeric(
      palette = c(
        "white", "orange"), 
      domain = c(0, 9)
    )
  )
```


## Creating a jTRACE-compatible language from the `phonemes` dataset

Imagine we wanted to run some simulations on a model that is familiar to the Spanish phonology. We could do so by shortlisting those phonemes present in Spanish (`is_spanish` is `TRUE`) and then following the steps in the [Creating a language](https://gongcastro.github.io/jtracer/articles/create-language.html) vignette. Please note that allophonic relations and duration scalars cannot be yet modified.

```{r}
spanish_phonemes <- phonemes %>% 
  filter(is_spanish) %>% 
  select(trace, pow:bur)

jtrace_create_language(
  language_name = "spanish",
  features = spanish_phonemes
)  

```

Now we are ready to launch jTRACE running `jtrace_launch()` and loading the `languages/spanish.xml` file in the *Phonemes* tab.

How are the seven feature dimensions related to each other?

```{r cors, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
library(ggcorrplot)

corr <- cor_pmat(phonemes[, 9:15])
colnames(corr) <- toupper(colnames(corr))
rownames(corr) <- toupper(colnames(corr))

ggcorrplot(
  corr, type = "upper", hc.order = TRUE, outline.color = "grey",
  lab = TRUE, colors = c("#f76560", "grey90", "#60f777"), show.diag = TRUE
) +
  theme_minimal() +
  theme(
    legend.position = "right",
    axis.title = element_blank(),
    axis.text = element_text(colour = "black"),
    panel.grid = element_blank()
  )
```


## Appendix

This is the complete list of implemented phonemes: 


```{r list}
phonemes %>% 
  rename_at(vars(pow:bur), toupper) %>%
  rename_at(vars(starts_with("is_")), ~gsub("is_", "", .)) %>% 
  gt(rowname_col = "id", groupname_col = "type") %>% 
  tab_spanner(md("**Acoustic features**"), columns = 6:12) %>% 
  cols_label(
    ipa = md("**IPA**"),
    trace = md("**jTRACE**"),
    description = ""
  ) %>% 
  data_color(
    columns = 6:12,
    colors = scales::col_numeric(
      palette = c(
        "white", "orange"), 
      domain = c(0, 9)
    )
  )
```


